Detection of Treatment-Subgroup Interactions in Clustered Datasets with Generalized Linear Mixed-effects Model Trees (GLMM trees)

### 2015-08-28 pkg/glmertree/paper/treatment_interactions.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@762 255f4249-2263-4417-b1ea-8498779e27a7

Although these findings are encouraging for the use of GLMM tree in the detection of treatment-subgroup interactions in datasets with clustered structures, some limitations should be noted.

The simulations show that GLMM tree performs very well, if the model is correctly specified. That is, if there are subgroups with respect to the partitioning variables, so that there are different parameters of the GLM in each of these subgroups, then the algorithm will accurately recover those subgroups. However, misspecification of the model can reduce performance. One source of misspecification would be, when relevant variables are not included in the GLM or as partitioning variables. If there are actual subgroups, but the variables describing them are not entered as partitioning variables, the algorithm can only approximate the subgroups using the partitioning variables that are available. Or, if the coefficients of other variables vary across subgroups, then those variables should also be included in the linear predictor of the GLM. Another source of misspecification would be the inclusion of irrelevant variables, either in the linear predictor of the GLM or as partitioning variables, which may reduce the power to detect the actual subgroups. However, it should be noted that in our simulations, the number of partitioning variables did not substantially influence performance of the algorithm(s).

>>>>

These findings are encouraging for the use of GLMM tree in the detection of treatment-subgroup interactions in datasets with clustered structures. However, it should be noted that the simulations show that GLMM tree performs very well, if the model is correctly specified. That is, if there are subgroups with respect to the partitioning variables, so that there are different parameters of the GLM in each of these subgroups, then the algorithm will accurately recover those subgroups. However, misspecification of the model can reduce performance. One source of misspecification would be, when relevant variables are not included in the GLM or as partitioning variables. If there are actual subgroups, but the variables describing them are not entered as partitioning variables, the algorithm can only approximate the subgroups using the partitioning variables that are available. Or, if the coefficients of other variables vary across subgroups, then those variables should also be included in the linear predictor of the GLM. Another source of misspecification would be the inclusion of irrelevant variables, either in the linear predictor of the GLM or as partitioning variables, which may reduce the power to detect the actual subgroups. However, it should be noted that in our simulations, the number of partitioning variables did not substantially influence performance of the algorithm(s).
